---
title: "p8105_hw2_yl4358"
author: "YI LIU"
date: "9/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

## Read and clean the Mr. Trash Wheel sheet

```{r}
mr_trash_wheel = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(
   sports_balls = round(sports_balls, digit = 0),
   sports_balls = as.integer(sports_balls)
  ) 
```

## Read and clean precipitation data for 2017 and 2018

```{r}
prep_2018 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  mutate(year = "2018") %>% 
  drop_na(month, total)

prep_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B15") %>%
  janitor::clean_names() %>% 
  mutate(year = "2017") %>% 
  drop_na(month, total)

prep_df = 
  bind_rows(prep_2017, prep_2018) %>% 
  mutate(
    month = month.name[month]
  ) 

```


* the number of observations in `mr_trash_wheel` dataset is `r length(pull(mr_trash_wheel, dumpster))`, and the the number of observations in `prep_df` dataset is `r length(pull(prep_df, month))`.

*Key variables are `dumpster`, `month` and `year`.

*Total precipitation in 2018 is `r sum(pull(filter(prep_df, year == 2018), total))` inches. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(mr_trash_wheel, year == 2017), sports_balls))`.


# Problem 2


## clean the data in pols-month.csv

```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month],
   president = ifelse(prez_gop == 0, "dem", "gop")
  ) %>% 
  select(-prez_gop, -prez_dem, -day) 

```

## clean the data in snp.csv

```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>% 
  arrange(year, month) %>% 
  select(year, month, everything()) %>% 
  mutate(
    month = as.numeric(month),
    month = month.name[month]
  ) %>% 
  select(-day)
```

## tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  mutate(
    year = as.character(year),
    month = recode(month, jan = month.name[1], feb = month.name[2], mar = month.name[3], apr = month.name[4],
                   may = month.name[5], jun = month.name[6], jul = month.name[7], aug = month.name[8], 
                   sep = month.name[9], oct = month.name[10], nov = month.name[11], dec = month.name[12])
  )
```

## join the datasets

```{r}
join_dataset = left_join(pols_month, snp_df, by = c("year", "month"))
final_dataset = left_join(join_dataset, unemployment_df, by = c("year", "month"))
```

* The `pols_month` dataset has `r nrow(pols_month)` rows and `r ncol(pols_month)` columns; The `snp_df` dataset has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns; The `unemployment_df` dataset has `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns;

* The `final_dataset` has `r nrow(final_dataset)` rows and `r ncol(final_dataset)` columns. The range of years is `r min(pull(final_dataset, year))` to `r max(pull(final_dataset, year))`. Key variables are `year` and `month`.

# Problem 3

## Load and tidy the data

```{r}
popular_baby_names = 
  read_csv("./data/Popular_Baby_Names.csv", col_types = "icccii") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name),
    ethnicity = recode(ethnicity,"asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic", "hisp" = "hispanic" )
  ) %>% 
  distinct()
```

## Produce a well-structures, reader-friendly table showing the rank in popularity of the name "Olivia" as a female baby name over time

```{r}
rank_oliva_female = 
  filter(popular_baby_names, childs_first_name == "olivia" & gender == "female") %>% 
  select(-count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
) %>% 
  select(-gender, -childs_first_name)
```

## Produce a similar table showing the most popular name among male children over time.

```{r}
popular_name_male = 
  filter(popular_baby_names, gender == "male" & rank == "1") %>% 
  select(-count) %>% 
  select(-rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% 
  select(-gender)
```

## produce a scatter plot

```{r}
spec_group = 
  filter(popular_baby_names, gender == "male" & ethnicity == "white non hispanic" & year_of_birth == 2016)

ggplot(spec_group, aes(x = rank, y = count)) +
  geom_point(aes(color = count)) + ggtitle("Scatterplot") 
```